---
title: "Customize Plot Appearance"
author: "Daniel Lüdecke"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Customize Plot Appearance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5, warning = FALSE, message = FALSE)
```


This vignette shows how the plots created by the `sjp.*` and `plot_model()` functions of the **sjPlot** package can be customized.

The examples refer to `plot_grpfrq()`, but most arguments are similar across all plotting function of the **sjPlot** package.

## Tweaking plot appearance

The base function to globally change theme option for all sjp-function is `set_theme()`. Except for geom-colors and geom-sizes, all theme-options can be set via this function. This new theme will be applied to all following plots created with the **sjPlot** package.

There are various arguments to change colors, sizes, angles etc. of labels. Following example show changes to colors, sizes, angles, geom-outlines and theme.

```{r}
# load libraries
library(sjPlot)  # for plotting
library(sjmisc)  # for sample data
library(ggplot2) # to access ggplot-themes

# load sample data set
data(efc)

set_theme(
  geom.outline.color = "antiquewhite4", 
  geom.outline.size = 1, 
  geom.label.size = 2,
  geom.label.color = "grey50",
  title.color = "red", 
  title.size = 1.5, 
  axis.angle.x = 45, 
  axis.textcolor = "blue", 
  base = theme_bw()
)

plot_grpfrq(
  efc$e42dep, 
  efc$e16sex, 
  title = NULL, 
  geom.colors = c("cadetblue", "coral"), 
  geom.size = 0.4
)
```

## Using the Color Brewer palettes

All plotting functions support the usage of the [Colorbrewer](http://colorbrewer2.org) palettes. To apply a color brewer palette, use specify the palette as `geom.colors`. Any valid color brewer palette is recognized automatically.

```{r}
# blank theme
set_theme(
  base = theme_blank(),
  axis.title.size = .9,
  axis.textsize = .9,
  legend.size = .7,
  legend.title.size = .8,
  geom.label.size = 3
)

plot_grpfrq(
  efc$e42dep, 
  efc$e15relat, 
  geom.colors = "PuRd", 
  show.values = FALSE
)
```

An overview of all supported color codes can be obtained with `display.brewer.all()` from the `RColorBrewer` package.

```{r, eval=FALSE}
library(RColorBrewer)
display.brewer.all()
```

## Plot with flipped coordinates

The plot's axes can be flipped using `coord.flip = TRUE`. If needed, labels can be placed inside the bars with the `vjust` or `hjust` arguments. In such cases, you might need to adjust the label colors with `geom.label.color = "white"`.

```{r}
set_theme(geom.label.color = "white", geom.label.size = 3)

# labels appear very large due to export metrics
plot_grpfrq(efc$e42dep, efc$e16sex, coord.flip = TRUE)
```

## Adding plot margins

Plots with no margins towards the axes may look strange to some people (not to me, though). To restore the ggplot-default behaviour, use the `expand.grid` argument:

```{r results='hide', echo=FALSE}
set_theme(
  axis.title.size = .9,
  axis.textsize = .9,
  legend.size = .7,
  legend.title.size = .8,
  geom.label.size = 3
)
```

```{r}
plot_grpfrq(efc$e42dep, efc$e16sex, expand.grid = TRUE)
```

## Theme options

You can use any pre-defined theme from ggplot, like `theme_bw()`, `theme_classic()` or `theme_minimal()` as default theme.

```{r}
set_theme(base = theme_light())
plot_frq(efc$e42dep)
```

## Pre-defined themes

There is a set of pre-defined themes from the sjPlot-package. See `?"sjPlot-themes"` for a complete list.

```{r}
library(sjmisc)
data(efc)
efc <- to_factor(efc, e42dep, c172code)
m <- lm(neg_c_7 ~ pos_v_4 + c12hour + e42dep + c172code, data = efc)

# reset theme
set_theme(base = theme_grey())

# forest plot of regression model
p <- plot_model(m)

# default theme
p
# pre-defined theme
p + theme_sjplot()
```

## Pre-defined scales

There is also a new scale for **ggplot**-objects, `scale_color_sjplot()` and `scale_fill_sjplot()`. 

```{r}
p + 
  theme_sjplot2() + 
  scale_color_sjplot("simply")
```

To see all currently available color sets, use `show_sjplot_pals()`.

```{r}
show_sjplot_pals()
```

## Set up own themes based on existing themes

If you want to use a specific theme as base for building your own theme, use the `base` argument. When using `base` instead of `theme`, further arguments for settings colors etc. are not ignored.
```{r}
set_theme(base = theme_bw(), axis.linecolor = "darkgreen")
plot_frq(efc$e42dep)
```


## Further customization options

Each plotting function invisibly returns the ggplot-object. You can further add options to customize the appearance of the plot, like in the following example, where the count axis is hidden (color set to white):

```{r}
set_theme(
  base = theme_classic(),
  axis.tickslen = 0, # hides tick marks
  axis.title.size = .9,
  axis.textsize = .9,
  legend.size = .7,
  legend.title.size = .8,
  geom.label.size = 3.5
)
  
plot_grpfrq(
  efc$e42dep,
  efc$e16sex,
  coord.flip = TRUE,
  show.axis.values = FALSE
) +
  theme(axis.line.x = element_line(color = "white"))
```

## Plot legend

The plot's legend can be customized via various `legend.`-arguments, see following examples:

```{r}
set_theme(
  base = theme_classic(), 
  legend.title.face = "italic", # title font face
  legend.inside = TRUE,         # legend inside plot
  legend.color = "grey50",      # legend label color
  legend.pos = "bottom right",  # legend position inside plot
  axis.title.size = .9,
  axis.textsize = .9,
  legend.size = .7,
  legend.title.size = .8,
  geom.label.size = 3
)

plot_grpfrq(efc$e42dep, efc$e16sex, coord.flip = TRUE)
```


```{r}
set_theme(
  base = theme_classic(), 
  axis.linecolor = "white",     # "remove" axis lines
  axis.textcolor.y = "darkred", # set axis label text only for y axis
  axis.tickslen = 0,            # "remove" tick marks
  legend.title.color = "red",   # legend title color
  legend.title.size = 2,        # legend title size
  legend.color = "green",       # legend label color
  legend.pos = "top",           # legend position above plot
  axis.title.size = .9,
  axis.textsize = .9,
  legend.size = .7,
  geom.label.size = 3
)

plot_grpfrq(efc$e42dep, efc$e16sex)
```

