---
title: "Customize Plot Appearance"
author: "Daniel Lüdecke"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Customize Plot Appearance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5, warning = FALSE, message = FALSE)
```


This vignette shows how the plots created by the `sjp.` functions of the sjPlot package can be customized.

The examples refer to the `sjp.grpfrq` function, but most arguments are similar across all plotting function of the **sjPlot** package.

Tweaking plot appearance
------------------------
The base function to globally change theme option for all sjp-function is `sjp.setTheme()`, and to use presets of themes, `set_theme()`. Except for geom-colors and geom-sizes, all theme-options can be set via this function. This new theme will be applied to all following plots created with the `sjPlot` package.

There are various arguments to change colors, sizes, angles etc. of labels. Following example show changes to colors, sizes, angles, geom-outlines and theme.
```{r}
# load libraries
library(sjPlot)  # for plotting
library(sjmisc)  # for sample data
library(ggplot2) # to access ggplot-themes
# load sample data set
data(efc)
sjp.setTheme(geom.outline.color = "antiquewhite4", 
             geom.outline.size = 1, 
             geom.label.size = 2,
             geom.label.color = "grey50",
             title.color = "red", 
             title.size = 1.5, 
             axis.angle.x = 45, 
             axis.textcolor = "blue", 
             base = theme_bw())
sjp.grpfrq(efc$e42dep, 
           efc$e16sex, 
           title = NULL, 
           geom.colors = c("cadetblue", "coral"), 
           geom.size = 0.4)
```

### Using the Color Brewer palettes
All plotting functions support the usage of the [Colorbrewer](http://colorbrewer2.org) palettes. To apply a color brewer palette, use specify the palette as `geom.colors`. Any valid color brewer palette is recognized automatically.

```{r}
# blank theme
set_theme(theme = "blank",
          axis.title.size = .9,
          axis.textsize = .9,
          legend.size = .7,
          legend.title.size = .8,
          geom.label.size = 3)
sjp.grpfrq(efc$e42dep, 
           efc$e15relat, 
           geom.colors = "PuRd", 
           show.values = FALSE)
```

An overview of all supported color codes can be obtained with `display.brewer.all()` from the `RColorBrewer` package.

```{r, eval=FALSE}
library(RColorBrewer)
display.brewer.all()
```

Plot with flipped coordinates
-----------------------------
The plot's axes can be flipped using `coord.flip = TRUE`. If needed, labels can be placed inside the bars with the `vjust` or `hjust` arguments. In such cases, you might need to adjust the label colors with `geom.label.color = "white"`.

```{r}
sjp.setTheme(geom.label.color = "white",
             geom.label.size = 3)
# labels appear very large due to export metrics
sjp.grpfrq(efc$e42dep, 
           efc$e16sex, 
           coord.flip = TRUE)
```

Adding plot margins
-------------------
Plots with no margins towards the axes may look strange to some people (not to me, though). To restore the ggplot-default behaviour, use the `expand.grid` argument:

```{r results='hide', echo=FALSE}
sjp.setTheme(axis.title.size = .9,
             axis.textsize = .9,
             legend.size = .7,
             legend.title.size = .8,
             geom.label.size = 3)
```

```{r}
sjp.grpfrq(efc$e42dep,
           efc$e16sex,
           expand.grid = TRUE)
```

Theme options
-------------
You can use any pre-defined theme from ggplot, like `theme_bw()`, `theme_classic()` or `theme_minimal()` as default theme. See the list of supported themes at the online-help `?sjp.setTheme` under argument _theme_.

```{r}
sjp.setTheme(base = theme_light())
sjp.frq(efc$e42dep)
```

Pre-defined themes
------------------

There is a set of pre-defined themes from the sjPlot-package that can be applied by calling `set_theme()`:

```{r}
set_theme("blank")
sjp.frq(efc$e42dep)

set_theme("scatter")
sjp.frq(efc$e42dep,
        geom.colors = "#2255aa")
```


Set up own themes based on existing themes
------------------------------------------

If you want to use a specific theme as base for building your own theme, use the `base` argument. When using `base` instead of `theme`, further arguments for settings colors etc. are not ignored.
```{r}
sjp.setTheme(base = theme_bw(),            # bw-base
             axis.linecolor = "darkgreen") # with green axes.
sjp.frq(efc$e42dep)
```


Further customization options
-----------------------------
Each plotting function invisibly returns the ggplot-object. You can further add options to customize the appearance of the plot, like in the following example, where the count axis is hidden (color set to white):
```{r}
sjp.setTheme(
  base = theme_classic(),
  axis.tickslen = 0, # hides tick marks
  axis.title.size = .9,
  axis.textsize = .9,
  legend.size = .7,
  legend.title.size = .8,
  geom.label.size = 3.5
)
  
sjp.grpfrq(
  efc$e42dep,
  efc$e16sex,
  coord.flip = TRUE,
  show.axis.values = FALSE,
  prnt.plot = FALSE
)$plot +
  theme(axis.line.x = element_line(color = "white"))
```

Note that the `prnt.plot` argument is set to `FALSE` in the above example, so the function call itself does not produce a plot (otherwise, two plots would be plotted).

Plot legend
-----------
The plot's legend can be customized via various `legend.`-arguments, see following examples:

```{r}
sjp.setTheme(base = theme_classic(), 
             legend.title.face = "italic", # title font face
             legend.inside = TRUE,         # legend inside plot
             legend.color = "grey50",      # legend label color
             legend.pos = "bottom right",  # legend position inside plot
             axis.title.size = .9,
             axis.textsize = .9,
             legend.size = .7,
             legend.title.size = .8,
             geom.label.size = 3)
sjp.grpfrq(efc$e42dep, 
           efc$e16sex, 
           coord.flip = TRUE)
```


```{r}
sjp.setTheme(base = theme_classic(), 
             axis.linecolor = "white",     # "remove" axis lines
             axis.textcolor.y = "darkred", # set axis label text only for y axis
             axis.tickslen = 0,            # "remove" tick marks
             legend.title.color = "red",   # legend title color
             legend.title.size = 2,        # legend title size
             legend.color = "green",       # legend label color
             legend.pos = "top",           # legend position above plot
             axis.title.size = .9,
             axis.textsize = .9,
             legend.size = .7,
             geom.label.size = 3)
sjp.grpfrq(efc$e42dep, efc$e16sex)
```

