---
title: "Plotting Estimates (Fixed Effects) of Regression Models"
author: "Daniel Lüdecke"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting Estimates (Fixed Effects) of Regression Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", dev = "png", fig.width = 7, fig.height = 5, message = FALSE, warning = FALSE)
```

This document describes how to plot estimates as forest plots (or dot whisker plots) of various regression models, using the `plot_model()` function. `plot_model()` is a generic plot-function, which accepts many model-objects, like `lm`, `glm`, `lme`, `lmerMod` etc.

`plot_model()` allows to create various plot tyes, which can be defined via the `type`-argument. The default is `type = "fe"`, which means that fixed effects (model coefficients) are plotted. For mixed effects models, only fixed effects are plotted by default as well.

```{r, results='hide'}
library(sjPlot)
library(sjmisc)
library(sjlabelled)
data(efc)
```


## Fitting a logistic regression model

First, we fit an example model, which will be used in the following examples. The examples work in the same way for any other model as well.

```{r results='hide'}
# create binary response
y <- ifelse(efc$neg_c_7 < median(na.omit(efc$neg_c_7)), 0, 1)

# create data frame for fitting model
df <- data.frame(
  y = to_factor(y),
  sex = to_factor(efc$c161sex),
  dep = to_factor(efc$e42dep),
  barthel = efc$barthtot,
  education = to_factor(efc$c172code)
)

# set variable label for response
set_label(df$y) <- "High Negative Impact"

# fit model
m1 <- glm(y ~., data = df, family = binomial(link = "logit"))

# set variable label for service usage
set_label(efc$tot_sc_e) <- "Total number of services used by carer"
```

## Plotting estimates of generalized linear models

The simplest function call is just passing the model object as argument. By default, estimates are sorted in descending order, with the highest effect at the top.

```{r}
plot_model(m1)
``` 

The "neutral" line, i.e. the vertical intercept that indicates no effect (x-axis position 1 for most glm's and position 0 for most linear models), is drawn slightly thicker than the other grid lines. You can change the line color with the `vline.color`-argument.

```{r}
plot_model(m1, vline.color = "red")
``` 

## Sorting estimates

By default, the odds ratios are sorted from highest to lowest value. You can also keep the order of predictors as they were introduced into the model if you set `sort.est` to `FALSE`.
```{r}
plot_model(m1, sort.est = FALSE)
``` 

Another way to sort estimates is to use the `order.terms`-argument. This is a numeric vector, indicating the order of estimates in the plot. In the summary, we see that "sex2" is the first term, followed by the three dependency-categories (position 2-4), the Barthel-Index (5) and two levels for intermediate and high level of education (6 and 7).

```{r}
summary(m1)
```

Now we want the educational levels (6 and 7) first, than gender (1), followed by dependency (2-4)and finally the Barthel-Index (5). Use this order as numeric vector for the `order.terms`-argument.

```{r}
plot_model(m1, order.terms = c(6, 7, 1, 2, 3, 4, 5))
``` 

## Estimates on the untransformed scale 

By default, `plot_model()` automatically exponentiates coefficients, if appropriate (e.g. for models with log or logit link). You can explicitley (dis-)allow this transformation using the `exponentiate`-argument.

```{r}
plot_model(m1, exponentiate = FALSE)
``` 

## Showing value labels

By default, just the dots and error bars are plotted. Use `show.values = TRUE` to show the value labels with the estimates values, and use `show.p = FALSE` to suppress the asterisks that indicate the significance level of the p-values.

```{r}
plot_model(m1, show.values = TRUE)
``` 

