---
title: "Basics of the sjt Table Functions"
author: "Daniel Lüdecke"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basics of the sjt Table Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```


This vignette describes some basic features of all `sjt` functions from the **sjPlot** package. These functions create HTML table outputs.


Printing HTML tables
--------------------

Before starting, we need to load the required packages and sample data.

```{r, message=FALSE, warning=FALSE}
# load package
library(sjPlot)
library(sjmisc)
# load sample data set.
data(efc)
```

All `sjt` functions create a HTML page with the data table. This table, by default, is opened in the viewer pane of your IDE (in case you're using an IDE which also supports the viewer pane, see parameter `useViewer` for details). If a viewer pane is not available, the created HTML output is saved as temporary file and opened in your default webbrowser. The temporary files are deleted after your R session ends.

Copying table output to office or word processors
-------------------------------------------------

### Export table as HTML file to open in word processors

You can save the HTML page as file for further usage by specifying the `file` parameter. The saved HTML file can be opened by word processors like [LibreOffice](http://www.libreoffice.org) or [Microsoft Office](http://office.microsoft.com).

### Drag and drop from browser or RStudio viewer pane

You can directly drag and drop a table from the RStudio viewer pane oder browser into your word processor. Simply select the complete table with your mouse and drag it into office.


Character encoding
------------------
In some cases, you may have to specify a character encoding in order to get proper labels in the HTML tables. If you face problems with correctly displaying labels, use the `encoding` parameter to change the character encoding. This value dependes on your region where you live. Following example, which works for western European countries, is the default behaviour of all `sjt`-functions:

```{r eval=FALSE}
# don't need to do this, because all sjt-functions
# use this code as default encoding-detection
if (.Platform$OS.type == "unix") 
  encoding <- "UTF-8" 
else 
  encoding <- "Windows-1252"
sjt.frq(efc$e15relat, encoding = encoding)
```

This example first detects your operating system and then chooses the associated character encoding, which is used in the HTML file. If this does not work for you, you have to use the `encoding` parameter.


Knitr integration of HTML tables
--------------------------------
```{r echo=FALSE, results='hide'}
rinline <- function(code){
  html <- '<code class="r">CODE</code>'
  sub("CODE", code, html, fixed = TRUE)
}
```

Most `sjt` function invisibly return a structure with following values:

* the web page style sheet `page.style`
* the web page content `page.content` 
* the complete html-output `output.complete`
* the html-table with inline-css for use with knitr `knitr`

`page.style` and `page.content` are described below in the CSS section. If you want to include a HTML table in your knitr document, simply use an inline code chunk. Be sure to use `no.output=TRUE`, because else your output would also be opened in a viewer pane or webbrowser.

`r rinline("sjt.frq(efc\\$e42dep, no.output=TRUE)\\$knitr")`

If you want to include both the source code and the output in your HTML document, you first need to write your code chunk, which should not be evaluated. Note that you don't need to use `no.output = TRUE` and `$knitr` in this chunk.

**Chunk 1**

    ```{r eval=FALSE}`r ''`
    sjt.frq(efc$e42dep)
    ```

Then repeat your code line, using the `no.output = TRUE` and `$knitr` parameters. This code chunk, however, should be evaluated:

**Chunk 2**

`r rinline("sjt.frq(efc\\$e42dep, no.output=TRUE)\\$knitr")`

If the above two code chunks 1 and 2 are compiled, the results look like this in your HTML document:

```{r eval=FALSE}
sjt.frq(efc$e42dep)
```

`r sjt.frq(efc$e42dep, no.output=TRUE)$knitr`


Customizing table output with the CSS parameter
-----------------------------------------------
The table output is in in [HTML format](http://en.wikipedia.org/wiki/HTML). The table style (visual appearance) is formatted using [Cascading Style Sheets](http://en.wikipedia.org/wiki/Cascading_Style_Sheets). If you are a bit familiar with these topics, you can easily customize the appearance of the table output.

Many table elements (header, row, column, cell, summary row, first row or column...) have CSS-class attributes, which can be used to change the table style. Since each `sjt` function has different table elements and thus different class attributes, you first need to know which styles can be customized.

The HTML and CSS code, which is described below, stem from this table output:
`r sjt.frq(efc$e42dep, no.output=TRUE)$knitr`

### Retrieving customizable styles
Each `sjt` function invisibly returns several values. The return value `page.style` contains the style information for the HTML table. You can print this style sheet to console using the base R `cat` function:
```{r eval=TRUE, warning=FALSE}
cat(sjt.frq(efc$e42dep, no.output = TRUE)$page.style)
```

The HTML code is obtained by using the `page.content` return value. Since the `sjt.frq` function allows to plot multiple tables at once, this function returns a list of HTML tables as `page.content.list`. The following code prints the HTML code of the table to the R console:

```{r eval=TRUE, warning=FALSE}
cat(sjt.frq(efc$e42dep, no.output = TRUE)$page.content.list[[1]])
# not that other sjt-functions don't return a content-list, but
# just '$page.content'...
```

Now you can see which table elements are associated with which CSS class attributes. If you compare the `page.style` with the related `page.content`, you see that not all style attributes are used:

* `.arc` is the attribute for the parameter `altr.row.col` and appears only in the `page.content` if it is set to TRUE.
* `.qrow` and `.mdrow` are the class attributes for the `emph.quart` and `emph.md` parameters.
* `.firsttablecol` is empty, but used. This class attribute is assigned in case you want to customize the appearance of the first table column (the one with value labels).

### Customizing table output with the CSS parameter
You can customize the table output with the `CSS` parameter. This parameter requires a list of attributes, which follow a certain pattern:

1. each attributes needs a `css.` prefix
2. followed by the class name (e.g. `caption`, `thead`, `centeralign`, `arc` etc.)
3. equal-sign
4. the CSS format (in (single) quotation marks)
5. the CSS format must end with a colon (;)

Example:
```{r eval=FALSE}
sjt.frq(efc$e42dep, 
        CSS = list(css.centeralign = 'text-align: left;', 
                   css.caption = 'font-weight: normal; font-style: italic;', 
                   css.firsttablecol = 'font-weight: bold;', 
                   css.lasttablerow = 'border-top: 1px solid; border-bottom: none;', 
                   css.summary = 'color: blue;'))
```
`r sjt.frq(efc$e42dep, no.output=TRUE, CSS=list(css.centeralign='text-align: left;', css.caption='font-weight: normal; font-style: italic;', css.firsttablecol='font-weight: bold;', css.lasttablerow='border-top: 1px solid; border-bottom: none;', css.summary='color: blue;'))$knitr`

In the above example, the summary-table row lost the original style and just became blue. If you want to keep the original style and just *add* additional style information, use the plus-sign (+) as initial character for the parameter attributes. In the following example, the summary row keeps its original style and is additionally printed in blue:
```{r eval=FALSE}
sjt.frq(efc$e42dep, CSS = list(css.summary = '+color: blue;'))
```
`r sjt.frq(efc$e42dep, no.output=TRUE, CSS=list(css.summary='+color: blue;'))$knitr`
