---
title: "Summary of Bayesian Models as HTML Table"
author: "Daniel Lüdecke"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
params:
    EVAL: !r identical(Sys.getenv("NOT_CRAN"), "true")
---

<!--
vignette: >
  %\VignetteIndexEntry{Summary of Bayesian Models as HTML Table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
-->

```{r echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  comment = "#>", 
  message = FALSE,
  eval = if (isTRUE(exists("params"))) params$EVAL else FALSE
)
```

This vignette shows examples for using `tab_model()` to create HTML tables for mixed models. Basically, `tab_model()` behaves in a very similar way for mixed models as for other, simple regression models, as shown [in this vignette](tab_model_estimates.html).

```{r, results='hide', message=FALSE, warning=FALSE}
# load required packages
library(sjPlot)
library(sjmisc)
library(brms)

# load sample datasets
data("efc")
efc <- to_factor(efc, e42dep, c172code, c161sex, e15relat)
zinb <- read.csv("http://stats.idre.ucla.edu/stat/data/fish.csv")

# fit two sample models
m1 <- brm(
  bf(count ~ child + camper + (1 | persons), 
     zi ~ child + camper),
  data = zinb,
  family = zero_inflated_poisson(),
  cores = 4,
  iter = 1000
)

f1 <- bf(neg_c_7 ~ e42dep + c12hour + c172code + (1 |ID| e15relat))
f2 <- bf(c12hour ~ c172code + (1 |ID| e15relat))
m2 <- brm(
  f1 + f2 + set_rescor(FALSE), 
  data = efc, 
  cores = 4,
  iter = 1000
)
```

## Bayesian models summaries as HTML table

For Bayesian regression models, some of the differences to the table output from [simple models](tab_model_estimates.html) or [mixed models](tab_mixed.html) of `tab_models()` are the use of _Highest Density Intervals_ instead of confidence intervals, the Bayes-R-squared values, and a different "point estimate" (which is, by default, the median from the posterior draws).

```{r}
tab_model(m1)
``` 

## Multivariate response models

For multivariate response models, like mediator-analysis-models, it is recommended to print just one model in the table, as each regression is displayed as own "model" in the output.

```{r}
tab_model(m2)
``` 

## Just show one HDI-column

To show just one HDI-column, use `show.hdi50 = FALSE`.

```{r}
tab_model(m2, show.hdi50 = FALSE)
``` 

## Mixing multivariate and univariate response models

When both multivariate and univariate response models are displayed in one table, a column _Response_ is added for the multivariate response model, to indicate the different outcomes.

```{r}
tab_model(m1, m2, show.hdi50 = F)
``` 
