---
title: "Summary of Mixed Models as HTML Table"
author: "Daniel Lüdecke"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Summary of Mixed Models as HTML Table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE)
```

This vignette shows examples for using `tab_model()` to create HTML tables for mixed models. Basically, `tab_model()` behaves in a very similar way for mixed models as for other, simple regression models, as shown [in this vignette](tab_model_estimates.html).

```{r, results='hide', message=FALSE, warning=FALSE}
# load required packages
library(sjPlot)
library(lme4)
library(nlme)
data("sleepstudy")
data("Orthodont")
```

## Mixed models summaries as HTML table

Unlike tables for [non-mixed models](tab_model_estimates.html), `tab_models()` adds additional information on the random effects to the table output for mixed models. You can hide these information with `show.icc = FALSE` and `show.re.var = FALSE`. Furthermore, the R-squared values are marginal and conditional R-squared statistics, based on _Nakagawa et al. 2017_.

```{r}
m1 <- lmer(distance ~ age + Sex + (1 | Subject), data = Orthodont)
m2 <- lmer(Reaction ~ Days + (1 + Days | Subject), data = sleepstudy)

tab_model(m1, m2)
``` 

The marginal R-squared considers only the variance of the fixed effects, while the conditional R-squared takes both the fixed and random effects into account.

The p-value is a simple approximation, based on the t-statistics and using the normal distribution function. A more precise p-value can be computed using `p.val = "kr"`. In this case, which only applies to linear mixed models, the computation of p-values is based on conditional F-tests with Kenward-Roger approximation for the degrees of freedom (using the using the **pbkrtest**-package). Note that here the computation is more time consuming and thus not used as default. You can also display the approximated degrees of freedom with `show.df`.

```{r}
tab_model(m1, p.val = "kr", show.df = TRUE)
``` 

## Generalized linear mixed models

`tab_model()` can also print and combine models with different link-functions.

```{r}
data("cvpp")
m3 <- glmer(
  cbind(incidence, size - incidence) ~ period + (1 | herd),
  data = cbpp, 
  family = binomial(link = "logit")
)

tab_model(m1, m3)
``` 

## More complex models

Finally, an example from the **glmmTMB**-package to show how easy it is to print zero-inflated generalized linear mixed models as HTML table.

```{r}
library(glmmTMB)
data("Salamanders")
m4 <- glmmTMB(
  count ~ spp + mined + (1 | site),
  ziformula = ~ spp + mined, 
  family = truncated_poisson(link = "log"), 
  data = Salamanders
)

tab_model(m1, m3, m4, show.ci = FALSE)
``` 

## References

Nakagawa S, Johnson P, Schielzeth H (2017) _The coefficient of determination R2 and intra-class correlation coefficient from generalized linear mixed-effects models revisted and expanded._ J. R. Soc. Interface 14. doi: 10.1098/rsif.2017.0213
